------------------------------------------------------------------------

------------------------------------------------------------------------

# Review of the study "Integrative single-cell analysis of transcriptional and epigenetic states in the human adult brain" by Lake, B., Chen, S., Sos, B. *et al.*

## Initializing libraries

```{r}
library(tidyverse)
library(SingleCellExperiment)
library(Matrix)
library(AnnotationHub)
library(ensembldb)
library(ggplot2)
library(scales)
library(Seurat)
```

## Creating Seurat Objects of all the samples

```{r}

counts <- read.table(file = "countmatrices/out_SRR5904942_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s42<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904943_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s43<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904944_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s44<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904945_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s45<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904946_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s46<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904947_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s47<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904948_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s48<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904949_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s49<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904950_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s50<- CreateSeuratObject(counts=counts, project = "Visual_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904951_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s51<- CreateSeuratObject(counts=counts, project = "Visual_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904952_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s52<- CreateSeuratObject(counts=counts, project = "Visual_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904953_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s53<- CreateSeuratObject(counts=counts, project = "Visual_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904954_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s54<- CreateSeuratObject(counts=counts, project = "Visual_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904955_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s55<- CreateSeuratObject(counts=counts, project = "Visual_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904956_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s56<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904957_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s57<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904958_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s58<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904959_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s59<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904960_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s60<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904961_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s61<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904962_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s62<- CreateSeuratObject(counts=counts, project = "Frontal_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904963_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s63<- CreateSeuratObject(counts=counts, project = "Frontal_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904964_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s64<- CreateSeuratObject(counts=counts, project = "Frontal_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904965_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s65<- CreateSeuratObject(counts=counts, project = "Frontal_Cortex")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904966_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s66<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904967_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s67<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904968_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s68<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)

counts <- read.table(file = "countmatrices/out_SRR5904969_gene_exon_tagged.dge.txt", header = TRUE, row.names = 1, colClasses =c("character", rep("numeric", 2500)))
s69<- CreateSeuratObject(counts=counts, project = "Cerebellar_Hemisphere")
rm(counts)
```

## Merging all the samples into a single Seurat Object

```{r}
so.merged <- merge(s42, y = c(s43, s44, s45, s46, s47, s48, s49, s50, s51, s52, s53, s54, s55, s56, s57, s58, s59, s60, s61, s62, s63, s64, s65, s66, s67, s68, s69), 
add.cell.ids = c("s42","s43","s44","s45","s46","s47","s48","s49","s50","s51","s52","s53","s54","s55","s56","s57","s58","s59","s60","s61","s62","s63","s64","s65","s66","s67","s68","s69"))

```

## Free individual samples memory

```{r}
rm(s42,s43,s44,s45,s46,s47,s48,s49,s50,s51,s52,s53,s54,s55,s56,s57,s58,s59,s60,s61,s62,s63,s64,s65,s66,s67,s68,s69)
```

# Quality Control

```{r}
mito.genes <- grep(pattern = "^MT-", x = rownames(so.merged@assays[["RNA"]]), value = TRUE)

percent.mito <- Matrix::colSums(so.merged@assays[["RNA"]][mito.genes, ])/Matrix::colSums(so.merged@assays[["RNA"]])

# AddMetaData adds columns to object@meta.data, and is a great place to
# stash QC stats

#Seurat v2 function, but shows compatibility in Seurat v3
so.merged <- AddMetaData(object = so.merged, metadata = percent.mito, col.name = "percent.mito") 
#in case the above function does not work simply do:
so.merged$percent.mito <- percent.mito

```

## Violin Plot

```{r}

VlnPlot(object = so.merged, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

```

```{r}

par(mfrow = c(1, 2))
FeatureScatter(object = so.merged, feature1 = "nCount_RNA", feature2 = "percent.mito")

```

```{r}

FeatureScatter(object = so.merged, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
```

```{r}

# We filter out cells that have unique gene counts (nFeature_RNA) over 2,500 or less than
# 200 Note that > and < are used to define a'gate'.  
#-Inf and Inf should be used if you don't want a lower or upper threshold.
so.merged <- subset(x = so.merged, subset = nFeature_RNA > 300 & nFeature_RNA < 5000 & percent.mito >  -Inf & percent.mito < 0.05 )
```

## Normalize the data

```{r}

VlnPlot(object = so.merged, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)

```

```{r}
so.merged <- NormalizeData(object = so.merged, normalization.method = "LogNormalize")
```

## Violin plot

```{r}
VlnPlot(so.merged, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
```

## Scatter Plot

```{r}
scatterplot <- FeatureScatter(so.merged, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
scatterplot
```

## Find variable genes

```{r}
rm(scatterplot)
so.merged <- FindVariableFeatures(so.merged, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(so.merged), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(so.merged)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1
```

## Scale data

```{r}
all.genes <- rownames(so.merged)
so.merged <- ScaleData(so.merged, vars.to.regress = c("nCounts_RNA", "percent.mito"))
```

```{r}
so.merged
```

## Run PCA

```{r}
so.merged <- RunPCA(object = so.merged, pc.genes = so.merged@var.genes, do.print = TRUE, pcs.print = 1:150, genes.print = 5)
```

## PCA Plot

```{r}
DimPlot(object = so.merged, reduction = "pca")

```

## Variable feature plot

```{r}
VariableFeaturePlot(object = so.merged)

```

## Heatmap

```{r}
DimHeatmap(object = so.merged, reduction = "pca", cells = 200, balanced = TRUE)
```

## Elbow Plot

```{r}
ElbowPlot(object = so.merged)
```

## Nearest Neighbors

```{r}
so.merged <- FindNeighbors(so.merged, reduction = "pca", dims = 1:15)
```

## Find clusters

```{r}
so.merged <- FindClusters(so.merged, resolution = 0.5, algorithm = 1)
```

## Run TSNE

```{r}
so.merged <- RunTSNE(object = so.merged, dims.use = 1:15, do.fast = TRUE, check_duplicates = FALSE)

```

```{r}
TSNEPlot(object = so.merged)
```

## TSNE by tissue type

```{r}
DimPlot(so.merged, reduction = "tsne", group.by = "orig.ident" )
```

## UMAP 20 PCs

```{r}
# Run UMAP map on first 20 PCs
so.merged <- RunUMAP(object = so.merged, dims = 1:20)
# Plot results
DimPlot(object = so.merged, reduction = 'umap')
```

## UMAP by tissue

```{r}
DimPlot(object = so.merged, reduction = 'umap', group.by = "orig.ident")
```

## Feature Plot

```{r}

FeaturePlot(object = so.merged, features = c("CHN1"), cols = c("grey", "blue"), reduction = "umap", pt.size = 0.1)
```

```{r}

FeaturePlot(object = so.merged, features = c("CREG2"), cols = c("grey", "blue"), reduction = "umap", pt.size = 0.1)


```

```{r}

FeaturePlot(object = so.merged, features = c("TFAP2B"), cols = c("grey", "blue"), reduction = "umap", pt.size = 0.1)

```

```{r}

FeaturePlot(object = so.merged, features = c("GRID2"), cols = c("grey", "blue"), reduction = "umap", pt.size = 0.1)
```
